Degrees Draft Syllabus

### 2023-03-21 Draft Syllabus.Rmd
continued updates to lesson 3

###### **select**
###### **rename**
###### **mutate**
###### **relocate**
###### **summarise**

```{r lesson3}

>>>>

###### **select**

Select is among the most common dplyr verbs. It has a simple, but important function: selecting which variables should be kept in a data frame. For example, if we only want to keep the mpg variable from the mtcars data frame, we can do the following:

```{r l3_2_ex13}
mtcars %>%
  select(mpg)
```

If instead, we'd like to keep multiple variables--say cylinders, gears, and carburetors:

```{r l3_2_ex14}
mtcars %>%
  select(mpg, gear, carb)

mtcars %>%
  select(c("mpg", "gear", "carb"))

mtcars %>%
  select(c(mpg, gear, carb))
```

As shown above, columns can be provided as a comma separated list, a character vector, or a vector of objects. This is especially useful if, instead, you want to remove some variables from the data frame:

```{r l3_2_ex15}
mtcars %>%
  select(-c("mpg", "gear", "carb"))
```

Now the variables of interest have been removed.

###### **rename**

Sometimes column names are inconvenient to work with, not descriptive, or generally do not meet your needs. Rename can remedy this problem; use it when you want to rename a column in a data frame:

```{r l3_2_ex16}
mtcars %>%
  rename(MilesPerGallon = mpg)
```

The new variable name is entered on the left-hand side of the equation, the old variable name on the right hand side. Multiple columns can be renamed at once by separating each expression with a comma:

```{r l3_2_ex17}
mtcars %>%
  rename(MilesPerGallon = mpg,
         Cylinders = cyl,
         Carburetors = carb)
```

R has rules about how variables can be named (see `help(make.names)` for more information). In general, variable names cannot have non-alphanumeric characters, cannot begin with numbers, and cannot contain spaces. Still, if you would really like to (for example if you are creating a public-facing table), you can create non-syntactic names by wrapping them in backticks:

```{r l3_2_ex18}
mtcars %>%
  rename(`Miles Per Gallon` = mpg)
```

###### **mutate**

Mutate allows you to modify and create columns in a data frame. You use mutate by writing the column name you'd like to modify/create on the left-hand side and an expression on the right-hand side of an expression. For example we can calculate the mpg to cyl ratio in mtcars:

```{r l3_2_ex19}
mtcars %>%
  mutate(mpg_to_cyl = mpg/cyl)
```

As with rename, if we'd like to manipulate multiple variables in one call to mutate we can do that by separating expressions with a comma:

```{r l3_2_ex20}
mtcars %>%
  mutate(mpg_to_cyl = mpg/cyl,
         mpg_to_carb = mpg/carb,
         carb_to_cyl = mpg_to_cyl/mpg_to_carb)
```

###### **relocate**
###### **summarise**
